\bf Faster Neural Machine Translation Inference

### 2018-02-22 fast-beam/fast-beam-short.tex
fast

As neural machine translation (NMT) models have become the new state-of-the-art, the challenge is to make their deployment efficient and economical. We have identified two areas where typical NMT models differ significantly from other deep-learning models that impact on their inference speed. Firstly, NMT models usually contains a large number of output classes, corresponding to the output vocabulary. Secondly, rather than a single label, the output from machine translation models is a sequence of class labels that makes up the words or sub-words in the target sentence. Target sentence lengths are unpredictable, leading to inefficiencies during batch processing. We provide solutions for each of these issues which together increase batched inference speed by up to 57\% on modern GPUs without affecting model quality.

@@

We examine two areas that are critical to fast NMT inference where the models differ significantly from other deep-learning models. We believe the optimizations of these models have been overlooked by the general deep-learning community.

Firstly, the number of classes in many deep-learning applications is small. However, the number of classes in NMT models is typically in the tens or hundreds of thousands, corresponding to the vocabulary size of the output language. For example, ~\citet{sennrich-haddow-birch:2016:P16-12} experimented with target vocabulary sizes of 60,000 and 90,000 sub-word units. This makes the output layer of NMT models very computationally expensive. Figure~\ref{fig:pie-time-europarl} shows the breakdown of amount of time during translation our NMT system; nearly 70\% of the time is involved in the output layer. We will look at optimizations which explicitly target the output layer.
\begin{figure}
\centering
\begin{tabular}{cc}
{\includegraphics[scale=0.3]{pie-time-europarl.png}} 
\end{tabular}
\caption{Proportion of time spent during translation (Europarl test set)}
\label{fig:pie-time-europarl}
\end{figure}

Secondly, the use of mini-batching is critical for fast model inference. However, mini-batching does not take into account variable sentence lengths which decrease the actual number of input or output sentences that are processed in parallel, negating the benefit of mini-batching. This can be partially reduced in the encoding with \emph{maxi-batching}, i.e. pre-sorting sentences by length before creating mini-batches with similar length source sentences. However, maxi-batching can introduce unacceptable delay in response for online applications as the input are not processed in the order they came. It is also not appropriate for applications where only the output lengths varies, for example, image captioning.

Even with sequence-to-sequence NMT models, target sentence lengths will still differ even for similar length inputs, compromising decoding performance. Figure~\ref{fig:batch-size} shows the actual batch size during decoding with a maximum batch size of 128; the batch was full in only 42\% of decoding iterations.

@@

We will propose an alternative batching algorithm to increase the actual batch size during decoding without the problems associated with mini-batching and maxi-batching.

\section{Prior Work}

Deep-learning models has been successfully used on many machine learning task in recent years in areas as diverse as computer vision and natural language processing. This success have been followed by the rush to put these model into production. The computational resources required in order to run these models has been challenging but, fortunately, there has been a lot work to meet this challenge.

Hardware accelerators such as GPUs are very popular but other specialized hardware such as custom processors~\citep{DBLP:journals/corr/JouppiYPPABBBBB17} and FPGA~\citep{DBLP:journals/corr/LaceyTA16} have been used. Hardware-supported reduced precision is also an ideal way to speed up model inference~\citep{DBLP:journals/corr/abs-1710-03740}.

Simpler, faster models have been created that are as good, or almost as good, as more slower, more complex models~\citep{DBLP:journals/corr/BahdanauCB14}. Research have also gone into smaller, faster models that can approximate slower, bigger models~\citep{DBLP:conf/emnlp/KimR16}. Some models have been justified as more suited for the parallel architecture of GPUs~\cite{DBLP:journals/corr/VaswaniSPUJGKP17}~\cite{gehring2017convs2s}.

The speed of the softmax layer when used with large vocabularies have been looked at in~\citet{DBLP:journals/corr/GraveJCGJ16} but there have been many other attempts at faster softmax, for example~\citet{DBLP:journals/corr/abs-1301-3781}, ~\citet{Zoph-2016}.

There has been surprisingly little work on batching algorithms, considering its critical importance for efficient deep-learning training and inference. ~\citet{Neubig-autobatching} describe an novel batching algorithm, however, its aim is to alleviate the burden on developers of batching models rather than faster batching.

Our paper follows most closely on from ~\citet{DBLP:conf/emnlp/Devlin17} which achieved faster NMT inference mainly by novel implementation of an existing model. However, we differ by focusing on GPU implementation, specifically the output layer, and the batching algorithm which was not touched on by the previous work. Our model is based on that of ~\citet{D14-1179} and ~\citet{sennrich-haddow-birch:2016:P16-12} but our work is applicable to other models and applications. % rather than CPUs as pursued by ~\cite{DBLP:conf/emnlp/Devlin17}.

@@

We leave step 1 for future work and focus on the last three steps, the outline for which are shown in Algorithm~\ref{algo:original-softmax-beamsearch}. For brevity, we show the algorithm for 1-best, a beam search of the n-bests is a simple extension of this.

@@

In fact, we are usually only interested in the best class during inference, not the probability. Since we now calculate the best class before the softmax, we can skip softmax altogether, Algorithm~\ref{algo:Argmax only}. However, this is only possible for beam size 1; the comparison between softmax probobilities is required for larger beam sizes as the denominator is different in different beams.

\begin{algorithm}
\begin{algorithmic}
\Procedure{Fused-kernel-1-best}{vector $p$, bias vector $b$}

\Return $best$

\end{algorithmic}
\caption{Find 1-best only}
\label{algo:Argmax only}
\end{algorithm}

@@

The standard mini-batching algorithm is outlined in Algorithm~\ref{algo:Mini-batching}.

@@

This algorithm encode the sentences for a batch, followed by decoding the batch. The decoding stop once all sentences in the batch are completed. This is a potential inefficiency as the number of remaining sentences may not be optimal.

We will focus on decoding as this is the more compute-intensive step, and issues with differing sentence sizes in encoding can partly be ameliorated by maxi-batching.

Our proposed top-up batching algorithm encode and decode asynchronously. The encoding step, Algorithm~\ref{algo:Encoding for top-up batching}, is similar to the main loop of the standard algorithm but the results are added to a queue to be consumed by the decoding step.

@@

Rather than decoding the same batch until all sentences in the batch are completed, the decoding step processing the same batch continuously. New sentences are added to the batch as old sentences completes, Algorithm~\ref{algo:Decoding for top-up batching}.

@@

We trained a sequence-to-sequence, encoder-decoder NMT system similar to that described in ~\citet{sennrich-haddow-birch:2016:P16-12}. This uses recurrent neural networks with gated recurrent units. The input and output vocabulary size were both set to 85,000 sub-words using byte-pair encoding (BPE) to adjust the vocabulary to the desired size. The hidden layer dimensions was set to 512. %We used the Marian toolkit to train our models.

@@

Our training data consisted of the German-English parallel sentences from the Europarl corpus~\citep{Koehn:2005:MTS}. To test inference speed, we used two test sets with differing characteristics:
\begin{enumerate}
   \item \vspace{-2 mm} a subset of the Europarl training data, which contains mostly long sentences, and is, of course, in the same domain as the training data
   \item \vspace{-2 mm} a subset of the German-English data from the Open-Subtitles corpus, consisting of mostly short, out-of-domain sentences.
\end{enumerate}
Table~\ref{tab:corpora} gives further details of the test sets.

\begin{table}
\begin{center}
\small
\begin{tabular}{|l|r|r|} \hline
		  & Europarl		& OpenSubtitles \\ \hline
\# sentences  	  & 30,000 		& 50,000 \\
\# sub-words 	  & 787,908 		& 467,654 \\ 
Avg sub-words/sent & 26.3		& 9.4 \\
Std dev subwords/sent & 14.9		& 6.1 \\ \hline
\end{tabular}
\end{center}
\caption{Test sets}
\label{tab:corpora}
\end{table}

@@

Fusing the last 3 steps led to a substantial reduction in the time taken, especially for beam size of 1 where the softmax probability does not actually have to calculated, Table~\ref{tab:fused-breakdown-europarl} and ~\ref{tab:fused-breakdown-opensubtitles}. This led to an overall increase in translation speed of up to 23\% for the Europarl test set, Figure~\ref{fig:beam-europarl}, and up to 41\% for the OpenSubtitles test set, Figure~\ref{fig:beam-opensubtitles}.

The amount of time taken by the output layer dominates translation time for the OpenSubtitles test set, Figure~\ref{fig:pie-time-opensubtitles}, explaining the bigger overall translation speed with the fused kernel.

\begin{table}
\begin{center}
\small
\begin{tabular}{|l|r|r|} \hline
		& Baseline	& Fused \\ \hline
\multicolumn{3}{|c|}{\emph{Beam size 1}}	\\ \hline	
Multiplication 	& 25.62 	& 26.47 (+3.3\%) \\ \hline
Add bias 	& 4.98		&  \\ 
Softmax 	& 12.81		& 7.94 (-76.9\%)\\
Beam search	& 16.64		&  \\ \hline
\multicolumn{3}{|c|}{\emph{Beam size 5}}	\\ \hline	
Multiplication 	& 112.9 	& 115.04 (+1.9\%) \\ \hline
Add bias 	& 23.66		&  \\ 
Softmax 	& 56.61		& 67.58 (-56.5\%)\\
Beam search	& 75.12		&  \\ \hline
\end{tabular}
\end{center}
\caption{Time taken in sec (Europarl)}
\label{tab:fused-breakdown-europarl}
\end{table}

@@

\begin{figure}
\centering
\begin{tabular}{cc}
{\includegraphics[scale=0.5]{beam-europarl.png}} 
\end{tabular}
\caption{Speed using the fused kernel (Europarl)}
\label{fig:beam-europarl}
\end{figure}

\begin{figure}
\centering
\begin{tabular}{cc}
{\includegraphics[scale=0.5]{beam-opensubtitles.png}} 
\end{tabular}
\caption{Speed using the fused kernel (OpenSubtitles)}
\label{fig:beam-opensubtitles}
\end{figure}

\begin{figure}
\centering
\begin{tabular}{cc}
{\includegraphics[scale=0.3]{pie-time-opensubtitles.png}} 
\end{tabular}
\caption{Proportion of time spent during translation (OpenSubtitles)}
\label{fig:pie-time-opensubtitles}
\end{figure}

@@

The top-up batching and maxi-batching have similar goals of maximizing efficiency when translating batches of different lengths. Therefore, using both methods together gives limited gains, in fact, using top-up batching with maxi-batch slows of between 2\% to 18\% when translating the Europarl test set due to the overhead of using the algorithm, Figure~\ref{fig:topup-europarl}. When maxi-batching is inappropriate, using top-up batching alone matches the performance of maxi-batching, even being slightly faster when a small beam is used.

\begin{figure}
\centering
\begin{tabular}{cc}
{\includegraphics[scale=0.3]{topup-europarl.png}} 
\end{tabular}
\caption{Speed using top-up batching (Europarl)}
\label{fig:topup-europarl}
\end{figure}

The results are better when translating the OpenSubtitles test set, Figure~\ref{fig:topup-opensubtitles}. The top-up batching does not harm performance when used with maxi-batching, even helping a little for small beams. However, top-up batching increases translation speed by up to 12\% when used alone.

\begin{figure}
\centering
\begin{tabular}{cc}
{\includegraphics[scale=0.3]{topup-opensubtitles.png}} 
\end{tabular}
\caption{Speed using top-up batching (OpenSubtitles)}
\label{fig:topup-opensubtitles}
\end{figure}

The gain from top-up batching is partially dependent on how much time the standard mini-batching algorithm spend decoding with a very small number of sentences as this is does not fully utilize the GPU cores. From Figure~\ref{fig:batch-size-small}, 20\% of the decoding iterations have less than 8 sentences remaining in the batch when translating the Europarl test test with maxi-batching. This figure is 50\% for the OpenSubtitles test set with no maxi-batching. The OpenSubtitles test set has a higher variance of sentence lengths relative to its average sentence length, which forces the mini-batch algorithm to continue decoding with a small number of sentences while the other sentences have already completed.

\begin{figure}
\centering
\begin{tabular}{cc}
{\includegraphics[scale=0.3]{batch-size-small.png}} 
\end{tabular}
\caption{Actual batch size during decoding}
\label{fig:batch-size-small}
\end{figure}

@@

Using both the fused kernel and top-up batching to translate led to a cummalative speed improvement of up to 57\% and 34\% for the OpenSubtitles and Europarl test set, respectively, when no maxi-batching is used, Figure~\ref{fig:cummulative-europarl} and Figure~\ref{fig:cummulative-opensubtitles}. With maxi-batching, the speed was up to 41\% and 21\%.

\begin{figure}
\centering
\begin{tabular}{cc}
{\includegraphics[scale=0.3]{cummulative-europarl.png}} 
\end{tabular}
\caption{Cummulative results (Europarl}
\label{fig:cummulative-europarl}
\end{figure}

\begin{figure}
\centering
\begin{tabular}{cc}
{\includegraphics[scale=0.3]{cummulative-opensubtitles.png}} 
\end{tabular}
\caption{Cummulative results (OpenSubtitles}
\label{fig:cummulative-opensubtitles}
\end{figure}

>>>>

We present two optimizations in areas where typical NMT models differ significantly from other deep-learning models. Firstly, we optimize the output layer which takes a significant amount of time for NMT models due to the large number of output classes, corresponding to the output vocabulary. Secondly, we present a novel batching algorithm which takes into account the differing output sentence lengths which leads to inefficiencies in current batch algorithms. Together, we increase inference speed by up to 57\% on modern GPUs without affecting model quality.

@@

The number of classes in NMT models is typically in the tens or hundreds of thousands, for example, ~\citet{sennrich-haddow-birch:2016:P16-12} experimented with target vocabulary sizes of 60,000 and 90,000 sub-word units. This makes the output layer of NMT models computationally expensive. Figure~\ref{fig:pie-time-opensubtitles} shows the breakdown of amount of time during translation our NMT system; 73\% of the time is involved in the output layer or beam search. Our first proposal will explicitly target this computation.
\begin{figure}
\centering
\begin{tabular}{cc}
{\includegraphics[scale=0.3]{pie-time-opensubtitles.png}} 
\end{tabular}
\caption{Proportion of time spent during translation}
\label{fig:pie-time-opensubtitles}
\end{figure}

Secondly, the use of mini-batching is critical for fast model inference. However, mini-batching does not take into account variable sentence lengths which decrease the actual number of input or output sentences that are processed in parallel, negating the benefit of mini-batching. This can be partially reduced in the encoding with \emph{maxi-batching}, i.e. pre-sorting sentences by length before creating mini-batches with similar length source sentences. However, maxi-batching can introduce unacceptable delays in response time as inputs are not processed in the order they came, a particular concern for online applications.

Even with sequence-to-sequence NMT models, target sentence lengths will still differ even for similar length inputs. Figure~\ref{fig:batch-size} shows the actual batch size during decoding with a maximum batch size of 128; the batch was full in only 42\% of decoding iterations. We will propose an alternative batching algorithm to increase the actual batch size during decoding without the problems associated with mini-batching and maxi-batching.

@@

% \section{Prior Work}
% 
% Deep-learning models has been successfully used on many machine learning task in recent years in areas as diverse as computer vision and natural language processing. This success have been followed by the rush to put these model into production. The computational resources required in order to run these models has been challenging but, fortunately, there has been a lot work to meet this challenge.
% 
% Hardware accelerators such as GPUs are very popular but other specialized hardware such as custom processors~\citep{DBLP:journals/corr/JouppiYPPABBBBB17} and FPGA~\citep{DBLP:journals/corr/LaceyTA16} have been used. Hardware-supported reduced precision is also an ideal way to speed up model inference~\citep{DBLP:journals/corr/abs-1710-03740}.
% 
% Simpler, faster models have been created that are as good, or almost as good, as more slower, more complex models~\citep{DBLP:journals/corr/BahdanauCB14}. Research have also gone into smaller, faster models that can approximate slower, bigger models~\citep{DBLP:conf/emnlp/KimR16}. Some models have been justified as more suited for the parallel architecture of GPUs~\cite{DBLP:journals/corr/VaswaniSPUJGKP17}~\cite{gehring2017convs2s}.
% 
% The speed of the softmax layer when used with large vocabularies have been looked at in~\citet{DBLP:journals/corr/GraveJCGJ16} but there have been many other attempts at faster softmax, for example~\citet{DBLP:journals/corr/abs-1301-3781}, ~\citet{Zoph-2016}.
% 
% There has been surprisingly little work on batching algorithms, considering its critical importance for efficient deep-learning training and inference. ~\citet{Neubig-autobatching} describe an novel batching algorithm, however, its aim is to alleviate the burden on developers of batching models rather than faster batching.
% 
% Our paper follows most closely on from ~\citet{DBLP:conf/emnlp/Devlin17} which achieved faster NMT inference mainly by novel implementation of an existing model. However, we differ by focusing on GPU implementation, specifically the output layer, and the batching algorithm which was not touched on by the previous work. Our model is based on that of ~\citet{D14-1179} and ~\citet{sennrich-haddow-birch:2016:P16-12} but our work is applicable to other models and applications. % rather than CPUs as pursued by ~\cite{DBLP:conf/emnlp/Devlin17}.

@@

We focus on the last three steps, the outline for which are shown in Algorithm~\ref{algo:original-softmax-beamsearch}. For brevity, we show the algorithm for 1-best, a beam search of the n-bests is a simple extension of this.

@@

In fact, we are usually only interested in the best class during inference, not the probability. Since we now compute the best class before the softmax, we can skip softmax altogether. This is only possible for beam size 1; the comparison between softmax probabilities is required for larger beam sizes as the denominators are different for different hypotheses.

@@

The standard mini-batching algorithm, Algorithm~\ref{algo:Mini-batching}, encode the sentences for a batch, followed by decoding the batch. Decoding stop only once all sentences in the batch are completed.

@@

Our proposal focuses on decoding as this is the more compute-intensive task, encoding and decoding asynchronously. The encoding step, Algorithm~\ref{algo:Encoding for top-up batching}, is similar to the main loop of the standard algorithm but the results are added to a queue to be consumed by the decoding step.

@@

Rather than decoding a batch until completion, new sentences are added to the batch as old sentences completes, Algorithm~\ref{algo:Decoding for top-up batching}.

@@

We train a sequence-to-sequence, encoder-decoder NMT system similar to that described in ~\citet{sennrich-haddow-birch:2016:P16-12}. This uses recurrent neural networks with gated recurrent units. The input and output vocabulary size are both set to 85,000 sub-words using byte-pair encoding (BPE), the hidden layer dimensions is 512. %We used the Marian toolkit to train our models.

@@

We trained a German-English with data from the Europarl corpus~\citep{Koehn:2005:MTS}, our test set consisted of a subset of the Open-Subtitles corpus, Table~\ref{tab:corpora}.

\begin{table}
\begin{center}
\small
\begin{tabular}{|l|r|r|} \hline
		 & OpenSubtitles \\ \hline
\# sentences  	 & 50,000 \\
\# sub-words 	 & 467,654 \\ 
Avg sub-words/sent & 9.4 \\
Std dev subwords/sent & 6.1 \\ \hline
\end{tabular}
\end{center}
\caption{Test sets}
\label{tab:corpora}
\end{table}

@@

Fusing the last 3 steps led to a substantial reduction in the time taken, especially for beam size of 1 where the softmax probability does not actually have to calculated, Table~\ref{tab:fused-breakdown-opensubtitles}. This led to an overall increase in translation speed of up to 41\%, Figure~\ref{fig:beam-opensubtitles}.

@@

\begin{figure}
\centering
\begin{tabular}{cc}
{\includegraphics[scale=0.5]{beam-opensubtitles.png}} 
\end{tabular}
\caption{Speed using the fused kernel}
\label{fig:beam-opensubtitles}
\end{figure}

@@

The top-up batching and maxi-batching have similar goals of maximizing efficiency when translating batches of different lengths. The top-up batching is of limited utility when used with maxi-batching but it increases translation speed by up to 12\% when used alone, Figure~\ref{fig:topup-opensubtitles}.

\begin{figure}
\centering
\begin{tabular}{cc}
{\includegraphics[scale=0.3]{topup-opensubtitles.png}} 
\end{tabular}
\caption{Speed using top-up batching}
\label{fig:topup-opensubtitles}
\end{figure}

@@

Using both the fused kernel and top-up batching to translate led to a cummalative speed improvement of up to 57\% when no maxi-batching is used, Figure~\ref{fig:cummulative-opensubtitles}. With maxi-batching, the speed was up to 41\% faster.

\begin{figure}
\centering
\begin{tabular}{cc}
{\includegraphics[scale=0.3]{cummulative-opensubtitles.png}} 
\end{tabular}
\caption{Cummulative results}
\label{fig:cummulative-opensubtitles}
\end{figure}

### 2018-02-22 fast-beam/fast-beam.tex
fast

In fact, we are usually only interested in the best class during inference, not the probability. Since we now calculate the best class before the softmax, we can skip softmax altogether, Algorithm~\ref{algo:Argmax only}. However, this is only possible for beam size 1; the comparison between softmax probobilities is required for larger beam sizes as the denominator is different in different beams.

>>>>

In fact, we are usually only interested in the best class during inference, not the probability. Since we now compute the best class before the softmax, we can skip softmax altogether, Algorithm~\ref{algo:Argmax only}. This is only possible for beam size 1; the comparison between softmax probabilities is required for larger beam sizes as the denominators are different for different hypotheses.
